\chapter{What is Security?}

\section{Overview}

The goal of this course is to give you an overview of the 
most important ``Big Ideas'' on securing computer systems.
Throughout the course, we will touch on ideas from the fields
of computer security, cryptography, and (to some extent)
computer systems.



\iffalse
\textbf{Big Idea}: Big ideas for securing computers.
\begin{itemize}
	\item Lectures: ask questions!
	\item Labs (coding) + psets (theory)
	\item Midterm + final
\end{itemize}
\fi

\section{What is Security?}
Security is a very broad property, but generally
the goal of computer security 
is to ensure that a particular computer system is 
\emph{behaves correctly} even in the
face of an \emph{adversary} (or \emph{attacker}) whose goal is to foil the
system.
\marginnote{We will use the terms ``adversary'' and ``attacker''
interchangeably throughout this course.}

To achieve this goal, we will need some kind of
systematic plan.
That is, we will have to carefully define 
what it means for our system to \emph{behave correctly}
and we will have to specify the class of 
\emph{adversaries} against which we want to defend.

For the purposes of this course, we will typically
structure our plan in terms of three components:
a \emph{model} of the system and the adversary,
a security \emph{goal}, 
an \emph{implementation}.

\begin{itemize}
\item \textbf{Model:}\marginnote{Sometimes people call this the ``threat model.''
    For the purposes of this chapter, think of the model as describing \emph{how} the 
    attacker and system interact and the security goal as defining \emph{what} our
    implementation is trying to achieve.
    }
    The system model specifies:
    \begin{enumerate}
      \item what the system is that we are trying to defend, 
      \item what the attacker is, and
      \item how the system and attacker interact.
    \end{enumerate}
    For example, in network security we might think of the
    system and the attacker as being two computers that
    interact over a network. In this model, 
    the attacker can send network packets
    to the victim system. (The attacker in this model cannot,
    for example, swap out the hard drive of the victim system.)

    When we are studying hardware security, we might consider
    a different model: The system we are trying to defend is a CPU,
    and the adversary is an external device that can 
    read all of the contents of the victim system's RAM. (The
    attacker in this model cannot, for example, read the internal
    state of the victim's CPU.)

    When we are studying ATM security, the system
    we are trying to defend is an ATM machine. For
    an attacker, we might consider a person
    interacting with the ATM through it's normal
    interface: the 
    attacker can insert a malformed ATM card into the machine and type
    arbitrary PINs into the device. (The attacker in this model cannot,
    for example, take a jackhammer to the ATM to extract the cash.)

    Modeling the adversary and the system in question is almost
    always the first step of thinking about system security.
    While we often only have an informal system model in mind, the more
    precise you---as a system designer---can be about your system model,
    the clearer your security properties can be.

  \item \textbf{Goal:} 
    The security goal defines what we want our system to achieve in our specified model.

    For example, in network security, we might want the property that ``only someone
    knowing Alice's secret password can execute shell commands on the machine.''
    In hardware security, we might want the property that ``the attacker learns nothing
    about the data stored in RAM, apart from its size.''
    In ATM security, we might want the property that ``an attacker can fraudulently 
    authenticate as a victim with probability at most $1/10000$.''

    As you will learn throughout the course, figuring out exactly what your
    security goal should be is often quite subtle and challenging.

  \item \textbf{Implementation:}
    The implementation is how we achieve the goal.
    For example, in securing a computer system on a network, we 
    might use password-based authentication to protect access to a computer system.
\end{itemize}

Together, the model and the goal model create our
\textit{definition} of security.\marginnote{When you read
about security failures in the news, it is worth trying
to understand whether the failure arose from a problem with the
model, the goal, or the implementation.}
As such, the model and goal
cannot be ``wrong''---%
the threat model might not have captured all of
the attacks that a real-world adversary can mount,
and goal might turn out to
not be exactly what we needed.
Often, the process of designing the model and goal is iterative:
when the system designer discovers a surprising gap in the security
goal (we will see some shortly), she patches the goal and the implementation
accordingly.

The implementation, on the other hand, can definitely be wrong---if 
the implementation does not guarantee the goal
under the model in question, due to bugs or
oversights or supply chain vulnerabilities or
anything else, the implementation has a mistake.

\paragraph{The implementation is always public: Kerckhoff's principle.}
Throughout this course, we will always assume that the attacker knows
the implementation of whatever security mechanism we are using.
The only thing that we keep secret from the adversary are the system's
secret keys.
This is known as \emph{Kerckhoff's principle}.
The logic behind this way of thinking is that: (1) it is often
relatively easy for the attacker to learn bits of information 
about the system design and (2) it is much easier to replace
a set of cryptographic keys (if the attacker learns them) than
to redesign the entire system from scratch.


\subsection{Security is Hard}

Building secure systems is challenging.
There are at least two broad reasons for this.

\paragraph{Secure systems must defend against worst-case behavior.}
First, a secure system must defend against
\emph{all possible attacks} within the scope of 
the system model.
In contrast, when we are just concerned about functionality or
correctness, we are often satisfied with a system 
that performs well for the cases that users care about.
\marginnote{Some developers do worry about correctness
of their system for all possible inputs and corner cases; this is
is indeed the mindset that is often necessary for security.}
In other words, security is concerned with behavior in
\emph{worst-case situations}, while correctness is often
about behavior in 
\emph{expected situations} (i.e., \emph{average-case situations}).

For example, suppose that you are a car manufacturer and you
want to test a car stereo functions correctly.
Testing that the stereo  works well \emph{on average} (i.e., in expected situations)
is easy: turn the stereo on and off 10,000 times, try playing some music
through it, and accept it as working if all of these checks pass.
Testing that the stereo works well \emph{in the worst case} is not as
easy: it is possible that if someone connects an adversarial USB device that sends
some specially crafted malicious packets to the
stereo, they can hijack the car and cause it to explode. But you will never
find these malicious packets by random testing---only by careful 
inspection.\marginnote{These attacks are actually possible in practice! See: \cite{koscher2010experimental}}
A secure system must defend against all possible attacks within the threat model;
being certain that a system satisfies this strong security property is a challenge.


\paragraph{An implementation can never defend against all possible threats.}
When we specify a system model, we delimit the set of adversaries 
against which our implementation must defend.
But real-world adversaries can behave in ways that are 
outside of our model and thereby violate our security goals.

For example, someone besides a TA might be able to
access the grades file for our class by:
\begin{itemize}
	\item finding a bug in the server software,
	\item breaking into a TA's office,
  \item compromising the TA's laptop,
	\item stealing the password to an administrator's account,
	\item tricking a TA into disclosing grades,
	\item breaking the server's cryptography,
	\item getting a job at the registrar and making herself a TA.
\end{itemize}

And the list never ends. Because our threat model cannot capture all
possible threats, \textbf{security is never perfect}.
There will essentially always be \textit{some} attacker than can break your system.

This is why we need a threat model: the threat
model defines what kinds of attacks we worry about
and which we decide are out of scope.

\subsection{Designing security goals and a threat model}
Specifying security goals and
a threat model is all about comparing
the cost of defending against an attack with the
cost of that attack if it were to happen.
It is almost always impractical to exactly calculate
these costs, but this framework is useful
conceptually.
Cheap defenses that block major
holes are likely to be worth implementing, but
defending against an esoteric RF side channel that
could leak unimportant information is likely not.

Building a threat model always requires iterating---you will not get it right on the first try.
There is likely to be some type of attack that you didn't consider at first that ends up being important.

\subsection{Designing an implementation}
We will focus largely in this class on techniques
that have a big payoff---methods of developing
software and tools to use that eliminate entire
classes of attacks (or make them much harder). 

\section{Examples}

We give a handful of examples of security failures arising from poor
choices of security goals or threat model.

\subsection{Insufficient Attack Models}

\subsubsection{Assuming specific strategy: CAPTCHA}
CAPTCHAs were designed to be expensive to
solve via automation, but easy for a human to read.
Indeed it might be expensive to build an optical-character recognition 
system for CAPTCHAs in general, but attackers who want
to bypass CAPTCHAs do not do this.
Instead, they set up computer centers in countries where the cost of
labor is cheap. Attackers then pay people working in these 
centers to solve CAPTCHAs.\autocite{motoyama:recaptchas}
The result is that it costs some fraction of
a cent to solve a CAPTCHA. 
The cost of solving a CAPTCHA is still non-zero,
but the cost is much lower than the system designers
may have intended.

\subsubsection{Assuming low limit on computational power: DES}
There used to be an encryption standard called DES that had $2^{56}$ possible keys.
At the time that it was designed, the U.S.~government standards agencies asserted that it was
secure against even powerful attackers,
but today a modern computer can try all $2^{56}$ keys at only modest cost.\marginnote{For
example, \url{https://crack.sh/} uses an array of FPGAs to provide a service that
exhaustively checks all possible keys.}
(Academic researchers even at the time of DES's
design understood that 56-bit keys were not large
enough to prevent exhaustive
cryptanalysis.\autocite{DH77})

Because of the weakness of DES against modern computers,
everyone using the standard had to upgrade their block ciphers.
For example, MIT had to switch from using DES for authentication 
to newer block ciphers with longer keys, such as AES.

\subsubsection{Assuming a secure out-of-band channel: Two-factor authentication (2FA) via SMS}
Many 2FA systems use a text message for authentication, but an attacker then
just needs to convince the clerk at the AT\&T store to give them a new SIM card for your phone number.
\marginnote{These attacks are often called ``SIM swapping'' or ``SIM hijacking''.}

\subsubsection{Assuming a correct compiler: Xcode}\label{sec:intro:xcode}\marginnote{This
prescient paper from 1984 anticipated the XcodeGhost attack: \cite{thompson1984reflections}}
iPhone apps are normally created and compiled on a developer's machine, sent to Apple's App Store, and sent to iPhones from there. iPhone apps are created using a tool called Xcode that is normally downloaded from Apple servers. However, Xcode is a big piece of software and for developers behind China's firewall, it was very slow to download. Someone within China set up a much faster mirror of Xcode, and lots of developers in China used the version of Xcode from this mirror. However, this mirror was not serving exactly Apple's version of Xcode---instead, it was serving a slightly modified version of Xcode that would inject some malicious code into every app that was compiled with it. This took a long time to detect.
\marginnote{The Wikipedia article on XcodeGhost, \url{https://en.wikipedia.org/wiki/XcodeGhost}, provides more details about this attack.}

\subsection{Insufficient security goal}
\subsubsection{Business-class airfare}
An airline tried to add value to their
business-class tickets by allowing ticket-holders to 
change the ticket (i.e., the departure date, origin, and destination)
at any time with no fee.
One customer realized that they could board the flight
\textit{then} change their ticket.
The customer could then take an unlimited number of
business-class flights for the price of one.

In this case, the airline's goal did not meet
their real needs---perhaps they needed to add
an additional goal of the form ``every time someone takes a flight,
we get paid.''

\subsubsection{Sarah Palin's Email}
Sarah Palin had a Yahoo email account, and Yahoo
used security questions for password reset---their
goal may have been something like ``no one can
reset a user's password unless they know all of
the answers to the user's security questions.''
(The security questions are typically things like
``What is your mother's maiden name?'')
As it turned out, it was possible to find the answer
to all of Palin's account-recovery security questions 
on the Internet.\autocite{palin}
Yahoo's implementation may have been perfect, but their
goal did not provide any meaningful security for certain users.

\subsubsection{Instruction Set Architecture (ISA) Specification}
When defining ISAs for processors, computer architects
thought it would be acceptable for each instruction
to take a different number of cycles to execute.
This had big benefits for performance and for
compatibility, but as we'll talk about later in
the semester, researchers have recently exploited
this timing variability to perform sophisticated
attacks on wide ranges of processors.\autocite{hill:spectre-meltdown}
Even if a processor's implementation faithfully implements
the specification, the specification itself
allows for certain types of timing side-channel attacks.

\subsubsection{Complicated access-control policies}
A school in Fairfax, Virginia used an online
course-management software with a somewhat complex
access-control structure: each teacher is in charge of some
class, each class has many students, and each
student has many files. Teachers cannot access
student's files, and there is also
a superintendent that has access to all the files.
Teachers are able to change their students'
passwords, and are able to add students to their
class. It turned out that a teacher could add the
superintendent as a student, change the
superintendent's password, and then access all
files via the superintendent's account.
While each of the access-control policies individually
sounds reasonable, together they lead to a bad outcome.


\subsubsection{Mat Honan's Gmail Account}
A journalist for wired named Mat Honan had his Gmail account compromised via a clever attack.\autocite{honan}
Honan had a Gmail account.
Gmail's reset-password feature avoided using security questions, and instead used a backup email account.
The attacker triggered the reset-password feature, which sent a reset link
to Honan's Apple email account.

The attacker then attempted to gain access to Honan's Apple email account.
The attacker triggered the reset-password feature on the Apple account.
Apple's reset-password feature, in turn, required Honan's address and the last four digits of the credit card number.
The attacker was able to find Honan's address publicly, but could not easily find his credit-card number.

The attacker found this credit-card information through Honan's Amazon account.
Amazon, which knew his credit card number, required a full credit-card in order to reset an account.
However, Amazon allowed buying something for a certain account without logging in, 
so long as you provide a new credit card number.
It also allowed \textit{saving} this new credit card number to the user's Amazon account.
So, the attacker made a purchase on Amazon with the attacker's credit card.
Then, the attacker saved their own credit card number to Honan's Amazon account.
Next, the attacker triggered Amazon's reset-password feature, and---using the credit-card
number saved to the account---were able to reset Honan's 
Amazon password and access his Amazon account.
The attacker was then able to see the last four
digits of Honan's real credit card within his
Amazon account, use that to reset his Apple mail
account, and then use that to reset Honan's Gmail
account.

Complex chains of systems like this can be very hard to reason about,
but these interactions ultimately are security-critical.

\subsection{Buggy implementations}
Bugs, misconfigurations, and other mistakes are
the most common cause of security issues. A rule
of thumb to keep in mind is that every 1000 lines
of code will have around one bug. This is a very
rough estimate, but the basic idea is that more
code will have more bugs. An effective strategy to
reduce security vulnerabilities is to reduce the
amount of code in your system.

\subsubsection{Missing Checks: iCloud}
Apple's iCloud performs many functions---email, calendar, storage, and Find my iPhone. Each of these had their own way of logging in, but across all of them a common goal was to limit the attacker's ability to guess a user's password. To do this, they added rate limiting to all the login interfaces, allowing something like only 10 login attempts per hour---but they forgot the Find my iPhone login interface.\autocite{icloud} Because this authentication code was duplicated all over the place, there were many places to remember to add this rate limiting, but the attacker only needed one weak login interface to brute-force a password. In general, avoiding this repition will make it much easier to build a secure system.

\subsubsection{Insecure Defaults}
When you set up a new service, they almost always come with some defaults to make the setup simpler. Wi-Fi routers come with default passwords, AWS S3 buckets come with default permissions, and so on. These defaults can be convenient, but they are very important to security because many people will forget or neglect to change the default. Because of this, the default becomes the way that the system operates. In order to build a secure system, it is important that the default is secure.



\section{What are the general principles for secure system design?}
\subsection{Threat Models and Goals}
\begin{itemize}
	\item Create simple, general goals.
	\item Avoid assumptions (such as ``no one else will be able to get a user's SIM card'') through better designs.
	\item Learn and iterate.
	\item Practice Defense in Depth: don't rely on one single defense for all your security---it is useful to use backup defenses to guard against bugs that will inevitably come up in one defense.
\end{itemize}

\subsection{Implementation}
\begin{itemize}
	\item A simpler system will lead to fewer problems.
	\item Factor out the security-critical part into a small separate system or piece of code (for example, hardware security keys).
	\item Reuse well-designed code, such as well-tested crypto libraries.
	\item Understand and test the corner cases.
\end{itemize}



